\LaTeXe\ guide for the Journal of Functional Programming

### 2019-04-14 papers/jfp/1-intro.tex
Minor revision to fit JFP style

\item We identify two key design choices that are typically deeply wired into
  any build system: \emph{the order in which tasks are
  built}~(\S\ref{sec-dependency-orderings}) and \emph{whether or not a
  task is (re-)built}~(\S\ref{sec-out-of-date}). These choices turn out to
  be orthogonal, which leads us to a new classification of the design
  space~(\S\ref{sec-design-space}).

>>>>

\item We identify two key design choices that are typically deeply wired into
  any build system: \emph{the order in which tasks are
  built}~(\S\ref{sec-dependency-orderings}) and \emph{whether or not a
  task is rebuilt}~(\S\ref{sec-out-of-date}). These choices turn out to
  be orthogonal, which leads us to a new classification of the design
  space~(\S\ref{sec-design-space}).

### 2019-04-14 papers/jfp/2-background.tex
Minor revision to fit JFP style

\begin{figure}[h]
\begin{subfigure}[b]{0.32\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/make-example.pdf}}
\caption{Task dependency graph}
\end{subfigure}
\begin{subfigure}[b]{0.32\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/make-example-full.pdf}}
\caption{Full rebuild}
\end{subfigure}
\begin{subfigure}[b]{0.32\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/make-example-partial.pdf}}
\caption{Partial rebuild}
\end{subfigure}
\vspace{-2mm}
\caption{A task dependency graph and two build scenarios. Input files are shown
in rectangles, intermediate and output files are shown in rounded rectangles.
Modified inputs and files that are rebuilt are highlighted.
\label{fig-make}}
\vspace{-2mm}
\end{figure}

If the user runs \Make specifying \cmd{main.exe} as the desired output, \Make
will build~\cmd{util.o}~and~\cmd{main.o}, in any order since these tasks are
independent, and then \cmd{main.exe}. If the user modifies \cmd{util.h} and runs
\Make again, it will perform a \emph{full rebuild}, because all three tasks
transitively depend on \cmd{util.h}, as illustrated in Fig.~\ref{fig-make}(b).
On the other hand, if the user modifies \cmd{main.c} then a \emph{partial
rebuild} is sufficient: \cmd{util.o} does not need to be rebuilt, since its
inputs have not changed, see Fig.~\ref{fig-make}(c). Note that if the dependency
graph is \emph{acyclic} then each task needs to be executed at most once. Cyclic
task dependencies are typically not allowed in build systems but there are rare
exceptions, see~\S\ref{sec-iterative-compute}.

@@

\begin{figure}
\begin{subfigure}[b]{0.90\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/shake-example.pdf}}
\caption{Dependency graph produced after the previous build.}
\end{subfigure}
\begin{subfigure}[b]{0.90\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/shake-example-rebuild.pdf}}
\caption{Since \cmd{docs.txt} was modified, we rebuild \cmd{release.txt} and
\cmd{release.tar}, discovering a new dependency.}
\end{subfigure}
\vspace{-2mm}
\caption{Dynamic dependencies example: create \cmd{README} and add it to the
list of release documents \cmd{docs.txt}.\label{fig-shake}}
\vspace{-4mm}
\end{figure}

@@

\begin{figure}
\vspace{-3mm}
\centerline{\includegraphics[scale=0.28]{fig/shake-example-cutoff.pdf}}
\vspace{-4mm}
\caption{An early cutoff example: if a comment is added to \cmd{main.c}, the
rebuild is stopped after detecting that \cmd{main.o} is unchanged, since this
indicates that \cmd{main.exe} and its dependents do not need to be
rebuilt.\label{fig-cutoff}}
\vspace{-4mm}
\end{figure}

@@

\vspace{-0.5mm}
\subsection{\Bazel: A Cloud Build System}\label{sec-background-bazel}
\vspace{-0.5mm}

@@

\begin{figure}
\vspace{-2mm}
\begin{subfigure}[b]{0.25\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/bazel-example-checkout.pdf}}
\vspace{-1mm}
\caption{Download source files}
\end{subfigure}
\begin{subfigure}[b]{0.40\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/bazel-example-build.pdf}}
\vspace{-1mm}
\caption{Build \cmd{main.exe}}
\end{subfigure}
\begin{subfigure}[b]{0.31\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/bazel-example-rebuild.pdf}}
\vspace{-1mm}
\caption{Modify \cmd{util.c} and rebuild}
\end{subfigure}
\vspace{-2.5mm}
\caption{A cloud build example: (a)~download sources, (b)~build \cmd{main.exe}
by downloading it from the cloud and skipping intermediate files (only their
hashes are needed), (c)~modify \cmd{util.c} and rebuild \cmd{main.exe}, which
requires building \cmd{util.o} (nobody has compiled \cmd{util.c} before) and
downloading \cmd{main.o} (it is needed for linking \cmd{main.exe}). File hashes
are shown in circles, and non-materialised intermediates in dashed rounded
rectangles.\label{fig-bazel}}
\vspace{-3mm}
\end{figure}

\begin{table}[h]
\vspace{-2mm}
\caption{Build system differences.\label{tab-summary}}
\vspace{-4mm}
% \smaller
\centering
\begin{tabular}{l||l|l||l|c|c|c}
\hline
$\!$Build system$\!$& Persistent build information & Scheduler   & Dependencies    & Minimal & Cutoff & Cloud$\!$\\\hline
$\!$\Make       $\!$& File modification times      & Topological & Static          & Yes     & No     & No   $\!$\\
$\!$\Excel      $\!$& Dirty cells, calc chain      & Restarting  & Dynamic         & No      & No     & No   $\!$\\
$\!$\Shake      $\!$& Previous dependency graph    & Suspending  & Dynamic         & Yes     & Yes    & No   $\!$\\
$\!$\Bazel      $\!$& Cloud cache, command history & Restarting  & Dynamic$^{(*)}$ & No      & Yes    & Yes  $\!$\\\hline
\end{tabular}
\vspace{2mm}
$^{(*)}$At present, user-defined build rules cannot have dynamic dependencies.
\vspace{-8mm}
\end{table}

\vspace{-1.5mm}
\subsection{Summary}\label{sec-background-summary}
\vspace{-0.5mm}

>>>>

\begin{figure}%[h]
\begin{subfigure}[b]{0.32\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/make-example.pdf}}
\caption{Task dependency graph}
\end{subfigure}
\begin{subfigure}[b]{0.32\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/make-example-full.pdf}}
\caption{Full rebuild}
\end{subfigure}
\begin{subfigure}[b]{0.32\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/make-example-partial.pdf}}
\caption{Partial rebuild}
\end{subfigure}
% \vspace{-2mm}
\caption{A task dependency graph and two build scenarios. Input files are shown
in rectangles, intermediate and output files are shown in rounded rectangles.
Modified inputs and files that are rebuilt are highlighted.
\label{fig-make}}
% \vspace{-2mm}
\end{figure}

If the user runs \Make specifying \cmd{main.exe} as the desired output, \Make
will build~\cmd{util.o} and~\cmd{main.o}, in any order since these tasks are
independent, and then \cmd{main.exe}. If the user modifies \cmd{util.h} and runs
\Make again, it will perform a \emph{full rebuild}, because all three tasks
transitively depend on \cmd{util.h}, as illustrated in Fig.~\ref{fig-make}(b).
On the other hand, if the user modifies \cmd{main.c} then a \emph{partial
rebuild} is sufficient: \cmd{util.o} does not need to be rebuilt, since its
inputs have not changed, see Fig.~\ref{fig-make}(c). Note that if the dependency
graph is \emph{acyclic} then each task needs to be executed at most once. Cyclic
task dependencies are typically not allowed in build systems but there are rare
exceptions, see~\S\ref{sec-iterative-compute}.

@@

\begin{figure}
\begin{subfigure}[b]{0.90\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/shake-example.pdf}}
\caption{Dependency graph produced after the previous build.}
\end{subfigure}
\begin{subfigure}[b]{0.90\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/shake-example-rebuild.pdf}}
\caption{Since \cmd{docs.txt} was modified, we rebuild \cmd{release.txt} and
\cmd{release.tar}, discovering a new dependency.}
\end{subfigure}
% \vspace{-2mm}
\caption{Dynamic dependencies example: create \cmd{README} and add it to the
list of release documents \cmd{docs.txt}.\label{fig-shake}}
% \vspace{-4mm}
\end{figure}

@@

\begin{figure}
% \vspace{-3mm}
\centerline{\includegraphics[scale=0.28]{fig/shake-example-cutoff.pdf}}
% \vspace{-4mm}
\caption{An early cutoff example: if a comment is added to \cmd{main.c}, the
rebuild is stopped after detecting that \cmd{main.o} is unchanged, since this
indicates that \cmd{main.exe} and its dependents do not need to be
rebuilt.\label{fig-cutoff}}
% \vspace{-4mm}
\end{figure}

@@

% \vspace{-0.5mm}
\subsection{\Bazel: A Cloud Build System}\label{sec-background-bazel}
% \vspace{-0.5mm}

@@

\begin{figure}
% \vspace{-2mm}
\begin{subfigure}[b]{0.25\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/bazel-example-checkout.pdf}}
% \vspace{-1mm}
\caption{Download sources}
\end{subfigure}
\begin{subfigure}[b]{0.40\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/bazel-example-build.pdf}}
% \vspace{-1mm}
\caption{Build \cmd{main.exe}}
\end{subfigure}
\begin{subfigure}[b]{0.31\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/bazel-example-rebuild.pdf}}
% \vspace{-1mm}
\caption{Modify \cmd{util.c} and rebuild}
\end{subfigure}
% \vspace{-2.5mm}
\caption{A cloud build example: (a)~download sources, (b)~build \cmd{main.exe}
by downloading it from the cloud and skipping intermediate files (only their
hashes are needed), (c)~modify \cmd{util.c} and rebuild \cmd{main.exe}, which
requires building \cmd{util.o} (nobody has compiled \cmd{util.c} before) and
downloading \cmd{main.o} (it is needed for linking \cmd{main.exe}). File hashes
are shown in circles, and non-materialised intermediates in dashed rounded
rectangles.\label{fig-bazel}}
% \vspace{-3mm}
\end{figure}

\begin{table}[h]
% \vspace{-2mm}
\caption{Build system differences.\label{tab-summary}}
% \vspace{-4mm}
% \smaller
\centering
\begin{tabular}{l||l|l||l|c|c|c}
\hline
$\!$Build system$\!$& Persistent build information & Scheduler   & Dependencies    & Minimal & Cutoff & Cloud$\!$\\\hline
$\!$\Make       $\!$& File modification times      & Topological & Static          & Yes     & No     & No   $\!$\\
$\!$\Excel      $\!$& Dirty cells, calc chain      & Restarting  & Dynamic         & No      & No     & No   $\!$\\
$\!$\Shake      $\!$& Previous dependency graph    & Suspending  & Dynamic         & Yes     & Yes    & No   $\!$\\
$\!$\Bazel      $\!$& Cloud cache, command history & Restarting  & Dynamic$^{(*)}$ & No      & Yes    & Yes  $\!$\\\hline
\end{tabular}
% \vspace{2mm}
$^{(*)}$At present, user-defined build rules cannot have dynamic dependencies.
% \vspace{-8mm}
\end{table}

% \vspace{-1.5mm}
\subsection{Summary}\label{sec-background-summary}
% \vspace{-0.5mm}

### 2019-04-14 papers/jfp/6-engineering.tex
Minor revision to fit JFP style

\begin{figure}
\begin{subfigure}[b]{0.32\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/frankenbuild-example-build.pdf}}
\vspace{-1.5mm}
\caption{Initial build}
\end{subfigure}
\begin{subfigure}[b]{0.32\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/frankenbuild-example-clean.pdf}}
\vspace{-1.5mm}
\caption{Clean up and evict \cmd{main.prof}}
\end{subfigure}
\begin{subfigure}[b]{0.34\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/frankenbuild-example-rebuild.pdf}}
\vspace{-1.5mm}
\caption{Build \cmd{main.prof} and \cmd{report.txt}}
\end{subfigure}
\vspace{-1.5mm}
\caption{A Frankenbuild example: (a)~build a human-readable profiling report for
\cmd{main.exe} from information dump \cmd{main.prof} produced by a profiling
tool, then record deep constructive traces in the cloud, (b)~remove built files
locally and evict \cmd{main.prof} from the cloud storage, (c)~build
\cmd{main.prof} by executing the profiler (profiling is non-deterministic, hence
the new hash value), then build \cmd{report.txt} by downloading it from the
matching deep constructive trace in the cloud, resulting in a Frankenbuild
because \cmd{main.prof} and \cmd{report.txt} are inconsistent. New and evicted
cloud storage entries are highlighted; file hashes are shown in circles.
\label{fig-frankenbuild}}
\vspace{-4mm}
\end{figure}

>>>>

\begin{figure}
\begin{subfigure}[b]{0.32\linewidth}
\centerline{\includegraphics[scale=0.26]{fig/frankenbuild-example-build.pdf}}
% \vspace{-1.5mm}
\caption{Initial build}
\end{subfigure}
\begin{subfigure}[b]{0.32\linewidth}
\centerline{\includegraphics[scale=0.26]{fig/frankenbuild-example-clean.pdf}}
% \vspace{-1.5mm}
\caption{Clean up, evict \cmd{main.prof}}
\end{subfigure}
\begin{subfigure}[b]{0.34\linewidth}
\centerline{\includegraphics[scale=0.26]{fig/frankenbuild-example-rebuild.pdf}}
% \vspace{-1.5mm}
\caption{Build \cmd{main.prof} \& \cmd{report.txt}}
\end{subfigure}
% \vspace{-1.5mm}
\caption{A Frankenbuild example: (a)~build a human-readable profiling report for
\cmd{main.exe} from information dump \cmd{main.prof} produced by a profiling
tool, then record deep constructive traces in the cloud, (b)~remove built files
locally and evict \cmd{main.prof} from the cloud storage, (c)~build
\cmd{main.prof} by executing the profiler (profiling is non-deterministic, hence
the new hash value), then build \cmd{report.txt} by downloading it from the
matching deep constructive trace in the cloud, resulting in a Frankenbuild
because \cmd{main.prof} and \cmd{report.txt} are inconsistent. New and evicted
cloud storage entries are highlighted; file hashes are shown in circles.
\label{fig-frankenbuild}}
% \vspace{-4mm}
\end{figure}
